Inverse design workflow of implicit surface-based cellular materials via conditional generative model and active Learning

### 2025-01-26 manuscript.tex
update on mac 2025-01-27

\begin{abstract}
Implicit surface-based cellular materials have gained widespread adoption in bone scaffolds applications due to their similarities to bone structure. These materials excel in providing good mechanical properties and bioadaptability, creating suitable conditions for bone tissue regeneration. However, the rational design of these materials typically depends on prior knowledge of experts and requires large of  trail-and-error experiments. Here, we present a data-efficient two-stage bi-objective optimization method within a machine learning framework that considers the varying costs associated different objectives. Specifically, we employ this methodology to orthopedic implant inverse design, integrating generative model, active leaning, and finite element analysis to achieve both biocompatible elastic modulus and higher yield strength cellular materials. Our approach demonstrates remarkable improvements in overall performance compared to initial datasets, while simultaneously reducing about sixfold design cost relative to one-step bi-objective optimization machine learning method. This methodology establishes an efficient paradigm for the rapid and intelligent design of cellular materials with tailored mechanical properties, particularly valuable in scenarios involving objectives with unbalanced computational costs.
\end{abstract}

@@

%%Research highlights
\begin{highlights}
\item Developing TPMS-GAN with navigator was proposed to generate TMPS unit cell under desired conditions of mechanical properties and geometries.
\item Developing Inversed TPMS-GAN to design TPMS unit cell in geomety feature in a explainable way instead of through latent space or coefficients of implicit function 
\item Proposing a constraint-aware active learning method to address bi-objective optimization problem with unbalanced costs of objective function. 
\end{highlights}

@@

With the rapid development of artificial intelligence (AI), machine learning has significantly transformed the design paradigm of cellular materials. By learning from data, machine learning models can effectively capture the non-linear relationships between geometric features and mechanical properties of cellular materials. Once trained, these models offer a low-cost inference capability, serving as surrogates for expensive numerical simulations and experiments. Consequently, machine learning models can be seamlessly integrated into heuristic algorithms facilitating the optimization of cellular material structures with target mechanical properties, such as elastic modulus \citep{Garland2021,Lee2022,Chen2024b}, bulk resistance \citep{Liu2020a}, and stress-strain curve \citep{Deng2022a,Ma2020b}.  However, surrogate optimization is a high-throughput selection process that necessitates multiple invocations of machine learning models, presenting challenges for global optimization problems and real-time design requirements.

@@

Figure \ref{fig:7}g indicates that the optimal structure was identified at the 2nd iteration, with a $Y$ value of 1.72, significantly higher than the optimal value of 0.22 in initial dataset. At the 3rd iteration, a decrease in overall performance was observed due to the active learning process, which integrates both exploration and exploitation capabilities. However, this approach led to improved performance at the 4th iteration. The stress-strain curve is depicted in Figure \ref{fig:7}i, while Figure \ref{fig:7}h presents the Von Mises' stress cloud plot. It is evident that stress concentration is localized at the junctions of unit cells. This study demonstrates that the constraint-aware active learning method successfully identified the optimal structure even beyond the training dataset, indicating a degree of extrapolation capability. Such robustness is particularly advantageous in clinical scenarios, where patient data, including target material properties and mechanical ranges, are often unknown in advance, and initial data distributions can vary widely. Furthermore, the design cost of TPMS unit cells with matched elastic modulus and high yield strength is significantly reduced by employing a method that requires the assessment of only 24 sample yield strengths. This is in stark contrast to the 1372 assessments needed in the one-step machine learning method used for training the generative model under bi-objective conditions. As a result, the design time is reduced about sixfold.

\begin{figure}
    \centering
    \includegraphics[width=1\linewidth]{figures/7.pdf}
    \caption{\textbf{Active learning workflow for yield stress of TPMS unit cells with elastic modulus constraint. a} Diagonal plot for SVM showing in red are the recommended structures and the green represents the initial data. \textbf{b} Performance of the SVM regressor as the iterations proceed. \textbf{c} Comparison of query points between EI and elastic modulus constraint-aware EI at 1st iteration. \textbf{d} Elastic modulus of the recommended structures as the iterations proceed. \textbf{e} The elastic modulus and yield strength distribution of recommended structures and the initial data (blue dots). \textbf{f} The elastic modulus and yield strength of recommended structures at each iteration of active learning. \textbf{g} The overall performance of TPMS unit cells as the iterations proceed. \textbf{h} The simulation strain-stress curves of the optimal structure recommended and the optimal 
structure in initial data. \textbf{g} The Von Mises' stress cloud plot of the optimal structure recommended in active learning.}
    \label{fig:7}
\end{figure}

@@

In summary, a data-efficient two-stage bi-objective inverse design method of cellular materials based on TPMS is proposed, which combines generative model, active learning and FEM. The approach considers the unbalanced costs associated with different objectives. The cubic symmetric TPMS unit cells (G and D) were megred to create the rich dataset of isotropic cellular material in which the effective elastic tensor are calculated using a voxel-based numerical homogenization method. Our proposed TPMG-GAN with a navigator can generate the TPMS unit cells under desired conditions, here we generate TPMS unit cells under a target elastic modulus. The inversed TPMS-GAN also developed to

% We proposed a GAN model to learn the IH mapping from properties to unit cell shapes that can be used to optimize functionally graded cellular structures. The cubic symmetric TPMS surfaces (P, D, and F-RD) were chosen and combined to create the dataset of isotropic cellular structures and represent each cellular unit cell as a six-dimensional shape vector (i.e., α1, α2, α3, t1, t2, t3). The unit cell property space consists of effective Young’s modulus, Poisson’s ratio, and relative density (i.e., E H , ν H , and ρ) in which E H and ν H were computed using a voxel-based numerical homogenization method. Our proposed IH-GAN’s one-to-many IH mapping was learned on the six-dimensional shape parameter space with the three-dimensional property space as the input conditions. Our approach offers an end-to-end generative model that automatically learns the IH mapping from data without assuming a bijective polynomial relationship. Except for the unit cell density (volume fraction), we also consider unit cell types when building the mapping. By including an auxiliary regressor in the IH-GAN, we can accurately generate the cellular unit cells that possess the desired properties (R2-scores between target properties and properties of generated unit cells > 98%). We also demonstrate the IH-GAN model’s efficacy by implementing it on modified structural optimization problems to construct multi-scale functionally graded cellular structures (e.g., a cantilever beam in this paper) using multiple types of cellular unit cells. Our approach addresses the connectivity issue between different types of unit cells simultaneously without a need for further compatibility optimization. By performing FE simulations, we validate that the beam constructed by IH-GAN improves the functional performance (e.g., 79.7% reduction in concentrated stress) compared to the conventional variable-density single-type beam structure. Moreover, the beam redesigned by IH-GAN can also achieve the target deformations with decent performance.

>>>>

\begin{abstract}
Implicit surface-based cellular materials have gained significant attention in bone scaffold applications due to their structural resemblance to natural bone. These materials exhibit excellent mechanical properties and bioadaptability, providing an ideal environment for bone tissue regeneration. However, the rational design of such materials often relies heavily on expert knowledge and extensive trial-and-error experiments. In the work, we propose a data-efficient, two-stage bi-objective optimization framework within a machine learning paradigm, which accounts for the disparate computational costs associated with different objectives. Specifically, we apply this methodology to the inverse design of orthopedic implants, integrating a generative model, active learning, and finite element analysis to achieve cellular materials with both biocompatible elastic modulus and enhanced yield strength. Our approach demonstrates significant improvements in overall performance compared to the initial dataset, while reducing design costs by approximately sixfold relative to conventional one-step bi-objective machine learning optimization methods. This framework establishes an efficient paradigm for the rapid and intelligent design of cellular materials with tailored mechanical properties, particularly in scenarios where objectives involve unbalanced computational costs.
\end{abstract}

@@

%%Research highlights
\begin{highlights}
\item A TPMS-GAN with a navigator was developed to generate TPMS unit cells under desired mechanical and geometric conditions.
\item An Inverse TPMS-GAN was introduced to design TPMS unit cells directly from geometric features in an explainable way, bypassingneed to interpolate in latent space or modify implicit function coefficients.
\item A constraint-aware active learning method was proposed to address bi-objective optimization problems with unbalanced objective function costs.
\end{highlights}

@@

With the rapid development of artificial intelligence (AI), machine learning has significantly transformed the design paradigm of cellular materials. By learning from data, machine learning models can effectively capture the non-linear relationships between geometric features and mechanical properties of cellular materials. Once trained, these models offer a low-cost inference capability, serving as surrogates for expensive numerical simulations and experiments. Consequently, machine learning models can be seamlessly integrated into heuristic algorithms facilitating the optimization of cellular material structures with target mechanical properties, such as elastic modulus \citep{Garland2021,Lee2022,Chen2024b}, bulk resistance \citep{Liu2020a}, and stress-strain curve \citep{Deng2022a,Ma2020b}.  However, surrogate optimization is a high-throughput selection process that requires multiple invocations of machine learning models, presenting challenges for global optimization problems and real-time design requirements.

@@

Figure \ref{fig:7}g indicates that the optimal structure was identified at the 2nd iteration, with a $Y$ value of 1.72, significantly higher than the optimal value of 0.22 in initial dataset. At the 3rd iteration, a decrease in overall performance was observed due to the active learning process, which integrates both exploration and exploitation capabilities. However, this approach led to improved performance at the 4th iteration. The stress-strain curve is depicted in Figure \ref{fig:7}i, while Figure \ref{fig:7}h presents the von Mises stress cloud plot. It is evident that stress concentration is localized at the junctions of unit cells. This study demonstrates that the constraint-aware active learning method successfully identified the optimal structure even beyond the training dataset, indicating a degree of extrapolation capability. Such robustness is particularly advantageous in clinical scenarios, where patient data, including target material properties and mechanical ranges, are often unknown in advance, and initial data distributions can vary widely. Furthermore, the design cost of TPMS unit cells with matched elastic modulus and high yield strength is significantly reduced by employing a method that requires the assessment of only 24 sample yield strengths. This is in stark contrast to the 1372 assessments needed in the one-step machine learning method used for training the generative model under bi-objective conditions. As a result, the design time is reduced about sixfold.

\begin{figure}
    \centering
    \includegraphics[width=1\linewidth]{figures/7.pdf}
    \caption{\textbf{Active learning workflow for yield stress of TPMS unit cells with elastic modulus constraint. a} Diagonal plot for SVM showing in red are the recommended structures and the green represents the initial data. \textbf{b} Performance of the SVM regressor as the iterations proceed. \textbf{c} Comparison of query points between EI and elastic modulus constraint-aware EI at 1st iteration. \textbf{d} Elastic modulus of the recommended structures as the iterations proceed. \textbf{e} The elastic modulus and yield strength distribution of recommended structures and the initial data (blue dots). \textbf{f} The elastic modulus and yield strength of recommended structures at each iteration of active learning. \textbf{g} The overall performance of TPMS unit cells as the iterations proceed. \textbf{h} The simulation strain-stress curves of the optimal structure recommended and the optimal 
structure in initial data. \textbf{g} The von Mises' stress cloud plot of the optimal structure recommended in active learning.}
    \label{fig:7}
\end{figure}

@@

In summary, this study proposes a data-efficient, two-stage bi-objective inverse design method for cellular materials based on TPMS. The method integrates a generative model, active learning, and finite element analysis, while accounting for the unbalanced costs associated with different objectives. Specifically, the methodology is applied to the inverse design of orthopedic implants. For TPMS-based bio-implants, two critical objectives are optimized: (1) the elastic modulus ($E$) of the scaffold implant must match that of the bone to mitigate the stress-shielding effect and promote bone recovery , and (2) the yield strength ($\sigma_{ys}$) must be maximized to ensure mechanical stability under load. Initially, cubic symmetric TPMS unit cells (G and D) were merged to create a comprehensive dataset of isotropic cellular materials, with the effective elastic tensor calculated using a voxel-based numerical homogenization method. The proposed TPMG-GAN, equipped with a navigator, generates TPMS unit cells under specified conditions, enabling the creation of a virtual structure space tailored to a target elastic modulus. Furthermore, an ITPMS-GAN was developed to design TPMS unit cell structures directly from geometric features in an explainable way, bypassing the need to interpolate in latent space or modify implicit function coefficients. This approach not only diversifies the virtual structure space but also provides a novel pathway for TPMS structure design. Finally, a constraint-aware active learning framework was extended to optimize yield strength under elastic modulus constraints, achieving a significant performance improvement from 0.2 to 1.7 compared to the initial dataset. This study demonstrates that the constraint-aware active learning method successfully identifies optimal structures even beyond the training dataset, showcasing its extrapolation capability. Additionally, the design time is reduced approximately sixfold compared to traditional one-step bi-objective machine learning methods.
